# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data_activity
```{r loaddata_activity}
Sys.setlocale("LC_ALL", "English_United States.1252")
Sys.setenv(LANG = "en_US.UTF-8")

unzip(zipfile="activity.zip")
data_activity <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
```{r}
library(ggplot2)
total.steps <- tapply(data_activity$steps, data_activity$date, FUN=sum, na.rm=TRUE)
qplot(total.steps, binwidth=1000, xlab= "number of steps/day")
mean(total.steps, na.rm=TRUE)
median(total.steps, na.rm=TRUE)
```

## What is the average daily activity pattern?
```{r}
library(ggplot2)
averages <- aggregate(x=list(steps=data_activity$steps), by=list(interval=data_activity$interval),
                      FUN=mean, na.rm=TRUE)
ggplot(data=averages, aes(x=interval, y=steps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("avg. number of steps")
```

On average across all the days in the data_activityset, the 5-minute interval contains
the maximum number of steps?
```{r}
averages[which.max(averages$steps),]
```

## Imputing missing values

There are many intervals with missing values. The presence of missing days may introduce bias into some calculations or summaries of the data_activity.

```{r how_many_missing}
missing <- is.na(data_activity$steps)
# How many missing
table(missing)
```

All of the missing values are filled in with mean value for that 5-minute interval.

```{r}
# Replace each missing value with the mean value of its 5-minute interval
fill.value <- function(steps, interval) {
    filled <- NA
    if (!is.na(steps))
        filled <- c(steps)
    else
        filled <- (averages[averages$interval==interval, "steps"])
    return(filled)
}
filled.data_activity <- data_activity
filled.data_activity$steps <- mapply(fill.value, filled.data_activity$steps, filled.data_activity$interval)
```
Using the filled data_activity set, let's make a histogram of the total number of steps taken each day and calculate the mean and median total number of steps.

```{r}
total.steps <- tapply(filled.data_activity$steps, filled.data_activity$date, FUN=sum)
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")
mean(total.steps)
median(total.steps)
```

Mean and median values are higher after the imputing step, because in the original data_activity dataset, there are some days with 'steps' values 'NA' for any 'interval.
By default those are set to 0s. Replacing them by the mean value increases the overall mean and median values.

## Are there differences in activity patterns between weekdays and weekends?
Let's find the day of the week for each measurement in the data_activityset. In
this part, we use the data_activity dataset with the filled-in values.

```{r}
weekday.or.weekend <- function(date) {
    day <- weekdays(date)
    if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
        return("weekday")
    else if (day %in% c("Saturday", "Sunday"))
        return("weekend")
    else
        stop("invalid date")
}
filled.data_activity$date <- as.Date(filled.data_activity$date)
filled.data_activity$day <- sapply(filled.data_activity$date, FUN=weekday.or.weekend)
```

Let's make a panel plot containing plots of average number of steps taken
on weekdays and weekends.

```{r}
averages <- aggregate(steps ~ interval + day, data=filled.data_activity, mean)
ggplot(averages, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) +
xlab("5-minute interval") + ylab("Number of steps")
```